A Large-scale Study of Polymorphism in R

### 2019-03-27 paper/main.tex
minor

In this section, we present our dataset. R aims to accommodate data
analysts, their workflows start with data import, followed by cleaning, and
then by steps of modeling, transformation and visualization. Often, the code
of these analysis pipelines resides, together with the data and results, in
notebooks. Few notebooks are publicly shared, and when they are, the data
isn't. For this reason our
analysis focuses on packages; these bundle reusable units of R code with
documentation, sample data and use-cases.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.15]{linesofrandccode.jpg}
\caption{Lines of code in the analyzed corpus. \# of lines of R code are above 0, \# of lines of C code are below 0.}
\label{allcloc}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.15]{recordsbypkg}
\caption{Numbers of Recorded Function Invocations in the Analyzed Corpus}
\label{recorded}
\end{center}
\end{figure}

@@

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:03:30 2019
\begin{table}[ht]
\label{tab:argcountsFGTS}
\centering
\begin{tabular}{lrr}
  \hline
 Argument Polymorphism & Count & Percentage \\ 
  \hline
  Full Monomorphic & 784K & 85.19 \\ 
  Monomorphic in Type & 839K & 91.27 \\ 
  Monomorphic in Class & 831K & 90.31 \\ 
  Monomorphic in Attribute Pattern & 830K & 90.25 \\ 
  Total Seen & 920K & --- \\ 
     \hline
\end{tabular}
\caption{Account of {\it argument} polymorphism in the FGTS.}
\end{table}

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:10:46 2019
\begin{table}[ht]
\label{tab:funcountsFGTS}
\centering
\begin{tabular}{lrr}
  \hline
 Function Polymorphism & Count & Percentage \\ 
  \hline
  Full Monomorphic & 143K & 65.64 \\ 
  Monomorphic in Type & 169K & 77.65 \\ 
  Monomorphic in Class & 163K & 75.13 \\ 
  Monomorphic in Attribute Pattern & 161K & 74.31 \\ 
  Total Seen & 217K & --- \\ 
   \hline
\end{tabular}
\caption{Account of {\it function} polymorphism in the FGTS.}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 21:27:37 2019
\begin{table}[ht]
\label{tab:toppolyFDTS}
\centering
\begin{tabular}{lrr}
  \hline
 Polymorphic Type Signature & \# Occurrences & \% of Total Polymorphism \\ 
  \hline
vector/double, vector/integer & 10,316 & 12.85 \\ 
scalar/double, scalar/integer & 7,194 & 8.96 \\ 
NULL, vector/double & 3,087 & 3.85 \\ 
data.frame, vector/double & 2,604 & 3.24 \\ 
NULL, scalar/character & 2,333 & 2.91 \\ 
list$<$any$>$, list$<$double$>$ & 2,224 & 2.77 \\ 
NULL, scalar/double & 2,066 & 2.57 \\ 
raw\_NA, vector/double & 1,465 & 1.83 \\ 
list$<$any$>$, list$<$list$>$ & 1,426 & 1.78 \\ 
NULL, vector/character & 1,385 & 1.73 \\ 
     \hline
\end{tabular}
\caption{Top polymorphic argument signatures in the FDTS.}
\end{table}

@@

The next type system design we will consider is ... .

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:03:30 2019
\begin{table}[ht]
\label{tab:argcountsWHATTS}
\centering
\begin{tabular}{lrr}
  \hline
 Argument Polymorphism & Count & Percentage \\ 
  \hline
  Full Monomorphic & 785K & 85.38 \\ 
  Monomorphic in Type & 874K & 95.08 \\ 
%  Monomorphic in Class & 831K & 90.31 \\ 
%  Monomorphic in Attribute Pattern & 830K & 90.25 \\ 
  Total Seen & 920K & --- \\ 
     \hline
\end{tabular}
\caption{Account of {\it argument} polymorphism in the NEW TYPE SYSTEM.}
\end{table}

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:10:46 2019
\begin{table}[ht]
\label{tab:funcountsWHATTS}
\centering
\begin{tabular}{lrr}
  \hline
 Function Polymorphism & Count & Percentage \\ 
  \hline
  Full Monomorphic & 143K & 65.95 \\ 
  Monomorphic in Type & 186K & 85.85 \\ 
%  Monomorphic in Class & 163K & 75.13 \\ 
%  Monomorphic in Attribute Pattern & 161K & 74.31 \\ 
  Total Seen & 217K & --- \\ 
   \hline
\end{tabular}
\caption{Account of {\it function} polymorphism in the NEW TYPE SYSTEM.}
\end{table}

We had some big wins here.

How do the top signatures shake out now?

>>>>

In this section, we present our dataset. R aims to accommodate data
analysts, their workflows start with data import, followed by cleaning, and
then by steps of modeling, transformation and visualization. Often, the code
of these analysis pipelines resides, together with the data and results, in
notebooks. Few notebooks are publicly shared, and when they are, the data
isn't. For this reason our analysis focuses on packages which bundle
reusable units of R code with documentation, sample data and use-cases.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.6]{linesofrandccode FILE MISSING IN GIT}
\caption{Lines of code in the analyzed corpus. \# of lines of R code are above 0, \# of lines of C code are below 0.}
\label{allcloc}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.1]{recordsbypkg}
\caption{Numbers of Recorded Function Invocations in the Analyzed Corpus}
\label{recorded}
\end{center}
\end{figure}

@@

In Figure~\ref{fig:matvec}, we see that the function can take in a vector, but immediately transforms it (transpose, with {\tt t}) into a matrix.

\begin{figure}[!hb]{\small\begin{lstlisting}[style=R]
tee <- function (x, theta, D1, D2, phi) 
{
    if (is.vector(x)) {
        x <- t(x)
    }
    ...
\end{lstlisting}}\caption{Example of matrix/(vector/double) argument (from {\tt calibrator} package).}\label{fig:matvec}\end{figure}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:03:30 2019
\begin{table}[ht]
\label{tab:argcountsFGTS}
\centering
\begin{tabular}{lrr}
  \hline
 Argument Polymorphism & Count & Percentage \\ 
  \hline
  Full Monomorphic & 784K & 85.28 \\ 
  Monomorphic in Type & 836K & 90.91 \\ 
  Monomorphic in Class & 831K & 90.31 \\ 
  Monomorphic in Attribute Pattern & 830K & 90.25 \\ 
  Total Seen & 920K & --- \\ 
     \hline
\end{tabular}
\caption{Account of {\it argument} polymorphism in the FGTS.}
\end{table}

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:10:46 2019
\begin{table}[ht]
\label{tab:funcountsFGTS}
\centering
\begin{tabular}{lrr}
  \hline
 Function Polymorphism & Count & Percentage \\ 
  \hline
  Full Monomorphic & 143K & 65.80 \\ 
  Monomorphic in Type & 167K & 76.84 \\ 
  Monomorphic in Class & 163K & 75.13 \\ 
  Monomorphic in Attribute Pattern & 161K & 74.31 \\ 
  Total Seen & 217K & --- \\ 
   \hline
\end{tabular}
\caption{Account of {\it function} polymorphism in the FGTS.}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 21:27:37 2019
\begin{table}[ht]
\label{tab:toppolyFDTS}
\centering
\begin{tabular}{lrr}
  \hline
 Polymorphic Type Signature & \# Occurrences & \% of Total Polymorphism \\ 
  \hline
  vector/double, vector/integer & 7677 & 9.18 \\ 
  scalar/double, scalar/integer & 7194 & 8.61 \\ 
  matrix, vector/double & 5354 & 6.40 \\ 
  NULL, scalar/character & 2333 & 2.79 \\ 
  list$<$any$>$, list$<$double$>$ & 2224 & 2.66 \\ 
  data.frame, matrix & 2115 & 2.53 \\ 
  NULL, scalar/double & 2066 & 2.47 \\ 
  NULL, vector/double & 2042 & 2.44 \\ 
  list$<$any$>$, list$<$list$>$ & 1426 & 1.71 \\ 
  NULL, vector/character & 1385 & 1.66 \\ 
     \hline
\end{tabular}
\caption{Top polymorphic argument signatures in the FDTS.}
\end{table}

@@

The next type system design we will consider is ... .
The number of monomorphic arguments and functions can be found in Tables~\ref{tab:argcountsWHATTS} and~\ref{tab:funcountsWHATTS} respectively.

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:03:30 2019
\begin{table}[ht]
\label{tab:argcountsWHATTS}
\centering
\begin{tabular}{lrr}
  \hline
 Argument Polymorphism & Count & Percentage \\ 
  \hline
  Full Monomorphic & 785K & 85.38 \\ 
  Monomorphic in Type & 871K & 94.65 \\ 
%  Monomorphic in Class & 831K & 90.31 \\ 
%  Monomorphic in Attribute Pattern & 830K & 90.25 \\ 
  Total Seen & 920K & --- \\ 
     \hline
\end{tabular}
\caption{Account of {\it argument} polymorphism in the NEW TYPE SYSTEM.}
\end{table}

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:10:46 2019
\begin{table}[ht]
\label{tab:funcountsWHATTS}
\centering
\begin{tabular}{lrr}
  \hline
 Function Polymorphism & Count & Percentage \\ 
  \hline
  Full Monomorphic & 143K & 65.95 \\ 
  Monomorphic in Type & 184K & 84.778 \\
%  Monomorphic in Class & 163K & 75.13 \\ 
%  Monomorphic in Attribute Pattern & 161K & 74.31 \\ 
  Total Seen & 217K & --- \\ 
   \hline
\end{tabular}
\caption{Account of {\it function} polymorphism in the NEW TYPE SYSTEM.}
\end{table}

We had some big wins here, with only ~5\% of argument polymorphism unaccounted for.
To get an idea of what more we can do, let's see what the most frequent remaining polymorphic signatures are.
See Table~\ref{tab:toppolyWHATTS}.

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Tue Mar 26 17:46:45 2019
\begin{table}[ht]
\label{tab:toppolyWHATTS}
\centering
\begin{tabular}{lrr}
  \hline
Polymorphic Type Signature & \# Occurrences & \% of Total Polymorphism \\ 
  \hline
  matrix, vector/real & 6307 & 12.82 \\ 
  list$<$any$>$, list$<$real$>$ & 2792 & 5.68 \\ 
  vector/character, vector/real & 2719 & 5.53 \\ 
  scalar/character, scalar/real & 2536 & 5.16 \\ 
  data.frame, matrix & 2163 & 4.40 \\ 
  list$<$any$>$, list$<$list$>$ & 1492 & 3.03 \\ 
  list$<$any$>$, list$<$character$>$ & 1217 & 2.47 \\ 
  data.frame, vector/real & 1025 & 2.08 \\ 
  scalar/logical, scalar/real & 917 & 1.86 \\ 
  scalar/character, vector/real & 844 & 1.72 \\ 
   \hline
\end{tabular}
\caption{Top 10 argument signatures in the NEW TYPE SYSTEM.}
\end{table}

What's up with matrix, vector/real?
